Predictive Modeling of I/O Characteristics \\ in High Performance Computing Systems

### 2017-11-11 HPC_Paper/paper.tex
Updated the algorithm comparison code to create the variance MDA data and also require an option saying which you want to predict. Will test local, then run on pawnee to collect variance prediction results.

Performance tuning is often an experimentally complex and time-intense
chore necessary for configuring HPC systems. The procedures for this
tuning vary largely from system to system and are often subjectively
guided by the system engineer(s). Once a desired level of performance
is achieved, HPC systems often remain in that base configuration with
all subsequent modifications being incremental. Future changes made
are often associated with updates and job-specific customizations. In
the case that a system has changing workloads or non-stationary
performance objectives that range from maximizing computational
throughput to minimizing power consumption and system variability, it
becomes obvious that a more effective and automated tool is needed for
configuring systems. This scenario presents a challenging and
important application of multivariate approximation and interpolation
techniques.

\begin{enumerate}
\item The value of multivariate Modeling
\item The data context
\item The proposed method for using multivariate models
\item The impact of effective models
\end{enumerate}

This paper compares five multivariate approximation techniques that
operate on inputs in $\mathbb{R}^d$ (vectors of $d$ real numbers) and
produce predicted responses in $\mathbb{R}^1$. In order to provide
coverage of the varied mathematical strategies that can be employed to
solve the continuous modeling problem, three of the techniques are
regression-based and the remaining two techniques produce
interpolants. The sections below outline the mathematical formulations
of each technique and provide computational complexity bounds.

\subsection{Approximation}
\subsubsection{Multivariate Adaptive Regression Splines}
\subsubsection{Multi-Layer Perceptron Regressor}
\subsubsection{Support Vector Regressor}

@@

%     Related Work     
%======================
\section{Related Work}
\begin{enumerate}
\item Not sure how much to include here? Shooting for thoroughness or
  simply necessary coverage? How much background should I expect the
  readers of this paper to have in the ``multivariate modeling of
  systems'' area?
\end{enumerate}

%     Methodology     
%=====================
\section{Methodology}
\subsection{Data}

\begin{table}
  \centering
  \begin{tabular}{|c|c|}

  \end{tabular}
  \caption{}
  \label{tab:data_type}
\end{table}

@@

@@

%     Future Work     
%=====================
\section{Future Work}

>>>>

Performance tuning is often an experimentally complex and time-intense
chore necessary for configuring HPC systems. The procedures for this
tuning vary largely from system to system and are often subjectively
guided by the system engineer(s). Once a desired level of performance
is achieved, an HPC system may only be incremenetally reconfigured as
required by updates or specific jobs. In the case that a system has
changing workloads or non-stationary performance objectives that range
from maximizing computational throughput to minimizing power
consumption and system variability, it becomes clear that a more
effective and automated tool is needed for configuring systems. This
scenario presents a challenging and important application of
multivariate approximation and interpolation techniques.

Predicting the performance of an HPC system is a challenging problem
that is primarily attempted in one of two ways: (1) build a
statistical model of the performance by running experiments on the
system at select settings or (2) use a performance simulator to run
artificial experiments and search for optimal configurations. In this
paper the proposed multivariate modelling techniques rest in the first
category, however they represent a notable increase in the ability to
model more complex interactions between system parameters.

Many previous works attempting to model system performance have used
simulated environments to estimate the performance of a
system. \cite{grobelny2007fase} \cite{wang2009simulation}
\cite{wang2013towards} Some of these works cite statistical models as
being over-simplified and not capabel of capturing the true complexity
of the underlying system. This claim is partially correct, noticing
that a large portion of predictive statistical models rely on
simplifying the system to one or two
parameters. \cite{snavely2002framework} \cite{bailey2005performance}
\cite{barker2009using} \cite{ye2010analyzing} These limited
statistical models have provided satisfactory performance in very
narrow application settings. Many of the aforementioned statistical
modeling techniques claim to generalize, while simultaneously
requiring additional code annotations, hardware abstractions, or
additional application level understandings in order to generate
models. The approach presented here requires no modifications of the
application, no architecture abstractions, nor any structural
descriptions of the input data being modelled. The techniques used are
purely mathematical and only need formatted data as input.

Among the statistcal models presented in prior works,
\cite{bailey2005performance} specifically mentioned that it is
difficult for the models to capture variability introduced by
I/O. System variability in general has become a problem of increasing
interest to the systems and HPC communities, however most of the work
has focused on Operating System induced variability.
\cite{beckman2008benchmarking} \cite{de2007identifying} The work that
has focused on managing I/O variability, \cite{lofstead2010managing},
does not use any sophisticated modeling techniques. Hence, this paper
presents a case study applying advanced mathematical and statistical
modeling techniques to the domain of HPC system I/O
characteristics. The models are used to predict the mean throughput of
a system and the variance in throughput of a system. The discussion
section outlines how the exact techniques presented can be applied to
any performance metric and any system.

%% Multivariate models for an HPC system would be a function of the
%% tunable parameters built to accurately model some desired performance
%% metric.

%% \begin{enumerate}
%% \item The value of multivariate Modeling
%% \item The data context
%% \item The proposed method for using multivariate models
%% \item The impact of effective models
%% \end{enumerate}

In general, this paper compares five multivariate approximation
techniques that operate on inputs in $\mathbb{R}^d$ (vectors of $d$
real numbers) and produce predicted responses in $\mathbb{R}^1$. In
order to provide coverage of the varied mathematical strategies that
can be employed to solve the continuous modeling problem, three of the
techniques are regression-based and the remaining two techniques
produce interpolants. The sections below outline the mathematical
formulations of each technique and provide computational complexity
bounds with respect to the size (number of points and dimension) of
input data.

\subsection{Approximation}
Multivariate approximations are capable of accurately modelling
complex relationships between the dimensions of any set of points in
$\mathbb{R}^n$ with respect to some response in $\mathbb{R}$. The
approximations produce a function $f: \mathbb{R}^n \rightarrow
\mathbb{R}$ which minimizes some metric related to the provided data.

\subsubsection{Multivariate Adaptive Regression Splines}
\subsubsection{Multi-Layer Perceptron Regressor}
\subsubsection{Support Vector Regressor}

@@

%% %     Related Work     
%% %======================
%% \section{Related Work}
%% \begin{enumerate}
%% \item Not sure how much to include here? Shooting for thoroughness or
%%   simply necessary coverage? How much background should I expect the
%%   readers of this paper to have in the ``multivariate modeling of
%%   systems'' area?
%% \end{enumerate}

%     Methodology     
%=====================
\section{Methodology}
\subsection{Data}
The summary of the data used in the experiments for this paper can be
seen in Table \ref{tab:data_type}.

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \textbf{System Parameter} & \textbf{Type} & \textbf{Values}\\
    \hline
    Hypervisor Scheduler & Categorical & CFQ, NOOP, DEAD\\
    Operating System Scheduler & Categorical & CFQ, NOOP, DEAD\\
    IOZone Test Type & Categorical & Fread, Fwrite, RandomRead\\
    File Size & Continuous & 64, 128, 256, 512, 1024\\
    Record Size & Continuous & 32 64, 128, 256, 512\\
    Thread Count & Continuous & 1,2,4,8,16,32,64\\
    Frequency & Continuous & 2.4, 2.5, 2.6, 2.7, 2.8, 2.9, 3.0, 3.1\\
    \hline
  \end{tabular}
  \caption{A description of the system parameters being considered in
    the IOZone tests. There are three categorical settings and four
    continuous settings. Notice that the ranges of values for
    continuous settings are different.}
  \label{tab:data_type}
\end{table}

@@

The framework used in this paper will be referred to as a
multi-dimensional analysis (MDA) of the IOZone data presented in this
study. However, the MDA framework can be applied to other datasets.

@@

%     Future Work     
%=====================
\section{Future Work}
The most severe limitation to the presented work is the present
inability to model the relationship between categorical system
parameters. This could be seen as a limitation of the models selected,
because only MARS is capable of handling non-numeric variables, or as
a limitation of the MDA framework.
